%
% This is your Bibliography file (bib.tex)
%
\begin{thebibliography}{99}
\bibitem{intro_miles_1980} Miles, J.W.,"Solitary waves", Ann. Rev. Fluid Mech., 12, 11-43
%\bibitem{intro_lee_1962}Lee, O.S., "Observations of internal waves in shallow water",
%\bibitem{intro_ziegenbein_1969}Ziegenbein, J., 1969 "Short internal waves in the Strait of Gibraltar", Deep Sea Res., 16, 479-487
%\bibitem{intro_ziegenbein_1970}Ziegenbein, J., 1970 "Spatial observations of short internal waves in the Strait of Gibraltar", Deep Sea Res., 17, 867-875
%\bibitem{intro_halpern_1971}Halpern, D., 1971, "Semidiurnal tides in Massachusetts Bay", J. Geophys. Res., 76, 6573-6584.
\bibitem{intro_lee_beardsley_1974}Lee, Ch.-Y., and R. C. Beardsley, 1974, "The Generation of long nonlinear internal waves in a weakly stratified shear flows", J. Geophys. Res., 79, n. 3, 453-457
\bibitem{intro_apel_1975a}Apel,. J.R., H. M. Byrne, J. R. Proni, and R.L. Charnell, 1975, "Observations of oceanic internal waves from the Earth Resources Technology Satellite", J. Geophys. Res., 80, 865-881
\bibitem{intro_ostr_step_1989}L. A. Ostrovsky and Y. A. Stepanyants, ÒInternal solitons in laboratory
experiments: Comparison with theoretical models,Ó Chaos 15, 037111 2005
\bibitem{intro_duda_farmer_1999}ÒThe 1998 WHOI/IOS/ONR Internal Solitary Wave Workshop: Contributed Papers,Ó edited by T. F. Duda and D. Farmer, Technical Report WHOI-99Ð07, 1999.
\bibitem{intro_apel_1995}J. R. Apel, ÒLinear and nonlinear internal waves in coastal and marginal
seas,Ó in Oceanographic Applications of Remote Sensing, edited by M. Ikeda and F. Dobson CRC Press, Boca Raton, 1995
\bibitem{intro_apel_1995} J. R. Apel, L. A. Ostrovsky, and Y. A. Stepanyants. "Internal solitions in the ocean". Technical Report MERCJRA0695, John Hopkins APL, July 1995.
\bibitem{intro_zhou_1991}Ji-xun Zhou, Xue-zhen Zhang, and Peter H. Rogers, "Resonant interaction of sound wave with internal solitons in the coastal zone," JASA 90(4), 2042-2054
\bibitem{intro_preisig_1997}J. C. Preisig and T. F. Duda, ÒCoupled acoustic mode propagation through continental shelf internal solitary waves,Ó IEEE J. Ocean. Eng. 222, 256-269 1997.
\bibitem{intro_Katznelson_2000}B. Katznelson and V. Petnikov, 2000,"Low frequency horizontal acousitc refraction caused by internal wave solitons in a shallow sea",{it\ Acoustical Physics}, 46 n.6, 774-788
\bibitem{intro_badiey_2002}M. Badiey, Y. Mu, J. F. Lynch, X. Tang, J. R. Apel, and S. Wolf, ÒAzimuthal and temporal dependence of sound propagation due to shallow water internal waves,Ó {\it IEEE J. Ocean. Eng.} 27(1), 117Ð129 2002.
\bibitem{intro_badiey_2005}M. Badiey, B. G. Katznelson, J. F. Lynch, S. A. Pereselkov, and W. Siegmann, ÒMeasurement and modeling of 3-D sound intensity variations due to shallow water internal waves,Ó J. Acoust. Soc. Am. 1172, 613Ð625 2005
\bibitem{intro_Rubenstein_1999}Rubenstein,D. 1999, "Observations of Cnoidal Internal Waves and Their Effect on Acoustic Propagation in Shallow Water ", {\it IEEE J. Ocean. Eng.}24(3), 346-357
\bibitem{intro_Lynch_2010}Lynch, J.F.; Ying-Tsong Lin; Duda, T.F.; Newhall, A.E. 2010,"Acoustic Ducting, Reflection, Refraction, and Dispersion by Curved Nonlinear Internal Waves in Shallow Water", {\it IEEE J. Ocean. Eng.} 35(1) 12-27
\bibitem{intro_Kropfli_1999}R. A. Kropfli,  L. A. Ostrovski,T. P. Stanton, E. A. Skirta, A. N. Keane,  and V. Irisov, "Relationships between strong internal waves in the coastal zone and their radar and radiometric signatures", {\it Journal of Geophysical Research}, Vol. 104, NO. C2, p.p. 3133-3148, Feb. 15, 1999 


\bibitem{1}   M. Badiey, Y. Mu, J. Lynch, J. R. Apel, and S.Wolf,
                "Temporal and azimuthal dependence of sound propagation
                in shallow water with internal waves," IEEE J. Oceanic Eng. 27(1), 117-129 (2002).
\bibitem{3}   M. Badiey, B. G. Katsnelson, J. F. Lynch, and S. Pereselkov, "Frequency
dependence and intensity fluctuations due to shallow water internal
waves," J. Acoust. Soc. Am. 122(2), 747-760 (2007).
\bibitem{jasa4}   Newhall, A. E., T. F. Duda, J. D. I. K. von der Heydt, J. N. Kemp,
S. A. Lerner, S. P. Liberatore, Y.-T. Lin, J. F. Lynch, A. R.
Maffei, A. K. Morozov, A. Shmelev, C. J. Sellers, and W. E. Witzell,
"Acoustic and oceanographic observations and configuration
information for the WHOI moorings for the SW06 experiment," WHOI
technical report WHOI-2007-04, (2007).
\bibitem{5}   Collis, J. M.,
T. F. Duda, J. F. Lynch, H. A. DeFerrari, "Observed limiting cases
of horizontal field coherence and array performance in a
time-varying internal wavefield," JASA, this issue (2008).

\bibitem{Tang1} D.J. Tang, J.N. Moum, J. Lynch, P. Abbott, R. Chapman, P. Dahl, T. Duda, G. Gawarkieweicz, S. Glenn, J. Goff, H. Graber, J. Kemp, A. Maffei, J. Nash and A. Newhall, "Shallow Water '06 A Joint Acoustic Propagation / Nonlinear Internal Wave Physics Experiment", Oceanography , 20(4), 156-167, 2007.
\bibitem{WHOI}Newhall, A.E., Duda, T.F., Von der Heydt, K., Irish, J.D., Kemp, J.N., Lerner, S., Liberatore, S.P, Lin, Y.T., Lynch, J.F., Maffei, A., Morozov, A.K., Shmelev, A., Sellers, C., and Witzell, W. ¡°Acoustic and oceanographic observations and configuration information for the WHOI moorings from the SW06 experiment,¡± Woods Hole. Oceanog. Inst. Tech Report, WHOI-2007-04, June 2007.

\end{thebibliography}
